MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-04-30 MAESTROeX_I/intro.tex
more updates

Previously, we developed the low Mach number astrophysical solver, MAESTRO.
Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
Furthermore, the low Mach number model in MAESTRO is specifically designed for astrophysical settings with significant atmospheric stratification.
This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
MAESTRO uses an explicit Godunov approach for advection, the stiff ODE VODE package \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
MAESTRO is a structured-grid, finite-volume code that utilizes adaptive mesh refinement (AMR) to refine grids locally in space.
This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.

@@

The rest of this paper is organized as follows.
In Section \ref{sec:equations} we review our model for stratified low Mach number astrophysical flow.
In Section \ref{eq:algorithm} we present our numerical algorithm in detail.
In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.
Finally, we conclude in Section \ref{sec:conclusions}.

>>>>

Previously, we developed the low Mach number astrophysical solver, MAESTRO.
Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited to, astrophysical settings with significant atmospheric stratification.
This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
MAESTRO uses an explicit Godunov approach for advection, the stiff ODE VODE package \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
MAESTRO is a structured-grid, finite-volume code that utilizes adaptive mesh refinement (AMR) to refine grids locally in space.
This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.

@@

The rest of this paper is organized as follows.
In Section \ref{sec:equations} we review our model for stratified low Mach number astrophysical flow.
In Section \ref{eq:algorithm} we present our numerical algorithm in detail, highlighting the new temporal integration scheme as well as spatial base state mapping options.
In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.
Finally, we conclude in Section \ref{sec:conclusions}.

### 2019-04-30 MAESTROeX_I/paper.tex
more updates

%% Mark off the abstract in the ``abstract'' environment.
\begin{abstract}
We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations allowing for improved scalability of the predecessor code.
We present a new, simplified temporal integration scheme that retains the same order of accuracy as our previous code.
We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
\end{abstract}

>>>>

%% Mark off the abstract in the ``abstract'' environment.
\begin{abstract}
We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
The underlying low Mach number equation set allows for efficient, long-time integration for highly subsonic flows compared to compressible approaches.
MAESTROeX is suitable for modeling full spherical stars as well as well as planar simulations of dynamics within localized regions of a star,
and can robustly handle several orders of magnitude of density and pressure stratification.
We present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches.
We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
The code leverages the new AMReX software framework for block-structured adaptive mesh refinement applications, allowing for scalability to large fractions of leadership-class machines.
Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae as a guide, we characterize the performance of the code and validate the new algorithmic features.
\end{abstract}
